---
title: "R Notebook"
output: html_notebook
---
Description
```{r}
?read.csv
```
Reads a file in table format and creates a data frame from it, with cases corresponding to lines and variables to fields in the file.

```{r}
?nrow
?ncol
```
nrow and ncol return the number of rows or columns present in x. NCOL and NROW do the same treating a vector as 1-column matrix, even a 0-length vector, compatibly with as.matrix() or cbind(), see the example.

```{r}
?colnames
```
Retrieve or set the row or column names of a matrix-like object.

```{r}
?sapply
```
lapply returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.

sapply is a user-friendly version and wrapper of lapply by default returning a vector, matrix or, if simplify = "array", an array if appropriate, by applying simplify2array(). sapply(x, f, simplify = FALSE, USE.NAMES = FALSE) is the same as lapply(x, f).

```{r}
?tapply
```
Apply a function to each cell of a ragged array, that is to each (non-empty) group of values given by a unique combination of the levels of certain factors.

```{r}
?sort
```
Sort (or order) a vector or factor (partially) into ascending or descending order. For ordering along more than one variable, e.g., for sorting data frames, see order.

```{r}
?sum
```
sum returns the sum of all the values present in its arguments
```{r}
?head
```
Returns the first or last parts of a vector, matrix, table, data frame or function. Since head() and tail() are generic functions, they may also have been extended to other classes.
```{r}
?length
```
Get or set the length of vectors (including lists) and factors, and of any other R object for which a method has been defined.
```{r}
?table
```
table uses the cross-classifying factors to build a contingency table of the counts at each combination of factor levels.
```{r}
?unique
```
unique returns a vector, data frame or array like x but with duplicate elements/rows removed.
```{r}
library(lubridate)
```
```{r}
?lubridate::mdy_hm
```
Transform dates stored as character or numeric vectors to POSIXct objects. The ymd_hms() family of functions recognizes all non-alphanumeric separators (with the exception of "." if frac = TRUE) and correctly handles heterogeneous date-time representations. For more flexibility in treatment of heterogeneous formats, see low level parser parse_date_time().
```{r}
?lubridate::date
```
Date-time must be a POSIXct, POSIXlt, Date, chron, yearmon, yearqtr, zoo, zooreg, timeDate, xts, its, ti, jul, timeSeries, and fts objects.
```{r}
?substring
```
Extract or replace substrings in a character vector.
```{r}
?mean
```
Generic function for the (trimmed) arithmetic mean.

Q1
```{r}
df<-read.csv("D:/TrainingData_latest.csv", header=T,sep=";",stringsAsFactors=F)
head(df)
```
Q2
```{r}
nrow(df)
ncol(df)
```
Q3
```{r}
Colname_df<-colnames(df)
Colname_df
```
Q4
```{r}
sapply(df,function(x){sum(is.na(x))})
```
Q5
```{r}
NA3<-sort(sapply(df,function(x){sum(is.na(x))}), decreasing = T)[1:3]
NA3
```
Q6
```{r}
LCID_name<-unique(df$LCID)
length(LCID_name)
```
Q7
```{r}
z<-mdy_hm(df$hour_id)
df$hour_id<-z
head(df)
```
Q8
```{r}
df$date<-date(df$hour_id)
tapply(df$hour_id,df$date, length)
```
```{r}
sum(tapply(df$hour_id,df$date, length))
```
Q9
```{r}
table(df$NumComplains)
```
Q10
```{r}
sort(tapply(df$NumComplains,df$LCID,sum),decreasing = T)[1:3]
```
Q11
```{r}
df$Region<-substring(df$LCID,1,5)
head(df$Region)
```
Q12
```{r}
sort(tapply(df$NumComplains,df$Region,sum),decreasing = T)[1:3]
```


